---
title: "Homework assignment 1"
subtitle: "Data Analysis 4: Prediction Analytics with Introduction to Machine Learning 2017/2018 Winter"
author: "Peter Paziczki"
date: '2018 február 11'
output: html_document
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Data Preparation

### 1.1 Loading the full London AirBnB dataset

I am loading the `airbnb_london_workfile.csv` data set, it has more 50,000 observations and 74 variables. Now I need to do proper data cleaning and preparation.

```{r echo=FALSE}
# loading necessary 
library(data.table)
library(ggplot2)

# Setting the path
path <- "/Users/User/Documents/R_projects/CEU-DA4/Seminar_1"
setwd(path)
#getwd()

# Loading the data
data <- fread(file = "airbnb_london_workfile.csv" , stringsAsFactors = F )

# Randomly choosing a borough
# data <- data[neighbourhood_cleansed =="Hackney"]
# Export the hackey data
#fwrite(data,"airbnb_hackney_workfile.csv")
```

### 1.2 Price

The target of this exercise is to predict price, so first let's have a better understanding of price variable.

```{r echo=FALSE}
summary(data$price)
# Create logs
data[,ln_price:=log(price)] # check for zeros
# Plot
qplot(data$price, geom="histogram", binwidth=50)+theme_bw()
qplot(data$ln_price, geom="histogram", binwidth=0.25)+theme_bw()
# Remove extreme values from prices
data <- data[!price>1000]
```

The mean and the median are far from each other and there seem to be a few very large numbers, the plot of price variable is skewed to the right. After having the log of price variable (`ln_price`) the there seems to be a more normal like distribution. I am dropping the observations that have a higher price than 1,000. The histograms of `price` and `ln_price` variable look better now in a sense of being less skewed are being closer to a normal like distribution, repsectively.

```{r}
# Much neater histograms
qplot(data$price, geom="histogram", binwidth=25, fill=I("lightblue"), col=I("white"))+theme_bw()
#ggsave("F14_h_price.png")

qplot(data$ln_price, geom="histogram", binwidth=0.15, fill=I("lightblue"), col=I("white"))+theme_bw()
#ggsave("F14_h_lnprice.png")
## it is closer to normal
```

### 1.3 Other variables

#### 1.3.1 Number of people accomodated

Let's have some summary statistics about prices by number of people accomodated.

```{r}
################################################
# look at some cnts. key vars, functional form #
################################################

## n_accomodates: look at distribution
data[,.(mean_price = mean(price) ,  min_price= min(price) ,max_price = max(price),  n=.N ),by = n_accommodates]

ggplot(data = data, aes(x=n_accommodates, y=price)) +
  geom_point(size=2, colour="orange")+
  ylim(0,800)+
  xlim(0,15)+
  labs(x="Number of people accomodated",y="Price")+
  geom_smooth(method="lm", colour="navy", se=FALSE)+
  theme_bw()
#ggsave("F14_s_n_accommodates.png")
## the higher number of people accomodated, the higher the price
```

There seems to be a general relation between `price` and `n_accomodates` variable,s, the more guest accomodated, the higher the price is on average.

```{r eval=FALSE}
# Squares and further values to create
data[, `:=`(n_accommodates2=n_accommodates**2, ln_accommodates=log(n_accommodates) ,
            ln_accommodates2=log(n_accommodates)**2) ]
# Regression 1: ln price and num of accomodates and squares
lm(ln_price ~ n_accommodates + n_accommodates2, data=data) # some kind of negative effect as number of accomodates increases
## let's see what would happen if we dropped the square ...

# Regression 2: ln price and log num of accomodates
lm(ln_price ~ ln_accommodates , data=data)
# Regression 3: ln price and num of accomodates
lm(ln_price ~ n_accommodates, data=data)

# lowess with scatterplot: price is lower than 800, num of acc.
ggplot(data = data[data$price<=800], aes(x=n_accommodates, y=price)) +
  geom_point(size=1.5, colour="orange", shape=4) +
  ylim(0,800)+
  xlim(0,18)+
  geom_smooth(method="loess", colour="darkgreen", se=F)+
  labs(x="Number of people accomodated",y="Daily price (USD)")+
  theme_bw()
#ggsave("F14_l_n_accommodates.png")
## there is some non-linearities

# lowess with scatterplot: log-price is lower than 800, log-num of acc.
ggplot(data = data[data$price<=800], aes(x=ln_accommodates, y=ln_price)) +
  geom_point(size=1.5, colour="orange", shape=4) +
  ylim(1,7)+
  xlim(0,3)+
  geom_smooth(method="loess", colour="darkgreen", se=F)+
  labs(x="Log number of people accomodated",y="Log daily price")+
  theme_bw()
#ggsave("F14_l_ln_accommodates.png")
# maybe best is to have log people -> better linear approximation, but different interpretation!!
## maybe we can say the using linears will be okay
## what are the warnings? when working with logs that are close to zero, there could be problems, but now everything is okay ...
```

#### 1.3.2 Beds

Let's have some summary statistics about prices by number of beds.

```{r}
## Beds
data[,.(mean_price = mean(price) ,  min_price= min(price) ,max_price = max(price),  n=.N ),by = n_beds]
# maybe best is to have log beds
data[,ln_beds:=log(n_beds)] # taking the log to have better interpretation
```

#### 1.3.3 Bathroom

Let's have a quick look at the histogram of bathrooms.

```{r}
## bathrooms
qplot(data$n_bathrooms, geom="histogram", binwidth=0.5, fill=I("lightblue"), col=I("white"))+theme_bw()
#ggsave("F14_h_n_bathrooms.png")
## there is a large number of accommodations with one bathroom
```

It would be wise to group the bathrooms into categorical variables based on their number per accommodation:

```{r}
# Pool accomodations with 0,1,2,10 bathrooms
data[,f_bathroom:=cut(n_bathrooms, c(0,1,2,10), labels=c(0,1,2), right = F)]
data[,.(mean_price = mean(price) ,  n=.N ),by = f_bathroom]
## creating factors to group accommodations based on the number of bathrooms
```

```{r}
## Number of reviews
qplot(data[data$n_number_of_reviews<100]$n_number_of_reviews,
      geom="histogram", binwidth=5, fill=I("lightblue"), col=I("white"))+ 
  labs(x="Number of reviews")+
  theme_bw()
# ggsave("F14_h_n_number_of_reviews.png")

# number of reviews: use logs as well
data[,ln_number_of_reviews:=log(n_number_of_reviews+1)]
qplot(data$ln_number_of_reviews,
      geom="histogram", binwidth=0.5, fill=I("lightblue"), col=I("white"))+ 
  theme_bw()
#ggsave("F14_h_ln_number_of_reviews.png")
## it is still exponential ... but not that high

# Pool num of reviews to 3 categories: none, 1-51 and >51
data[,f_number_of_reviews:=cut(n_number_of_reviews, c(0,1,51,max(data$n_number_of_reviews)), labels=c(0,1,2), right = F)]
data[,.(median_price = median(price) ,mean_price = mean(price) ,  n=.N ),by = f_number_of_reviews]
## grouping / pooling the reviews ... they are factors already ... median and mean are close to each other

# Regression 1: log-price and number of reviews
lm(ln_price ~ f_number_of_reviews, data=data) # might be a relevant variable ...
# Regression 2: log-price and log number of reviews
lm(ln_price ~ ln_number_of_reviews, data=data) # there seems to be a not so strong connection

## Time since
# Create variables, measuring the time since: squared, cubic, logs
data[,`:=`(ln_days_since= log(n_days_since),ln_days_since2 = log(n_days_since)**2
           ,ln_days_since3 = log(n_days_since)**3 , n_days_since2=n_days_since**2, n_days_since3=n_days_since**3)]
# Check the effect
ggplot(data = data[(data$price<=800) & (ln_days_since>2)], aes(x=ln_days_since , y=ln_price)) +
  geom_point(size=1.5, colour="orange", shape=4) +
  ylim(1,7)+
  xlim(2,7)+
  geom_smooth(method="loess", colour="darkgreen", se=F)+
  labs(x="Log number of days since first review",y="Log daily price")+
  theme_bw()
#ggsave("F14_l_ln_days_ince.png")
## there might be some pattern

#-Inf values
#lm(ln_price ~ ln_days_since + ln_days_since2 + ln_days_since3, data=data)

## review score effect
ggplot(data = data[(data$price<=800) & (n_review_scores_rating>=60)], aes(x=n_review_scores_rating , y=ln_price)) +
  geom_point(size=1.5, colour="orange", shape=4) +
  ylim(1,7)+
  xlim(60,100)+
  geom_smooth(method="loess", colour="darkgreen", se=F)+
  labs(x="Review score",y="Log daily price")+
  theme_bw()
#ggsave("F14_l_n_number_of_reviews.png")
## it can be meaningful and useful to not only look at the full population but subsamples too

# Create log of review scores
data[,ln_review_scores_rating :=log(n_review_scores_rating )]
# Regression 1) ln price - num of review scores
lm(ln_price ~ n_review_scores_rating,data=data) 
# Regression 2) ln price - log num of review scores
lm(ln_price ~ ln_review_scores_rating,data=data) 
#leave as is

## minimum nights
lm(ln_price ~ n_minimum_nights,data=data) # it is almost zero, let's pool the nights and see if it is more meaningful that way

# Pool and categorize the number of minimum nights: 1,2,3, 3+
data[,f_minimum_nights:=cut(n_minimum_nights, c(1,2,3,max(data$n_minimum_nights)), labels=c(1,2,3), right = F)]

lm(ln_price ~ f_minimum_nights,data=data)
## comparing to one night stay if minimum stay is two nights, then the percentage in price with one extra night is going to be larger by 24%

###########################
## look at categoricals  ##
###########################

categoricals <- c("f_property_type", "f_room_type", "f_cancellation_policy", "f_bed_type") # defines already defined as factors

for (var in categoricals) {
  print(data[,.(mean_price = mean(price) ,  n=.N ),by = mget(var)])
}

#####################################

# fwrite(data,"airbnb_hackney_workfile_adj.csv")

```